Report on the Third Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE3)

### 2015-10-23 WSSSPE3/report/WSSSPE3_review.tex
removing catalog group from README and main document
Section 3: keynote: More discussion on productization  and volunteer models.

Turk then moved onto various different definitions of sustainability. His 
favorite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding of new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software.
A funder Turk heard talked about sustainability as continuing to get funded. Turk also
recalled that Greg Wilson said in WSSSPE 1.1 that his view of sustainable software
was software that continued to give the same results over time. A last measure
of sustainability Turk presented was the ability to transition between different people
developing and using a piece of software.

@@

\input{wg_main_best_practices}
\input{wg_main_funding_spec_expert}
\input{wg_main_catalogs}
\input{wg_main_industry_interaction}
\input{wg_main_legacy_SW}
\input{wg_main_eng_design}
\input{wg_main_metrics}
\input{wg_main_training}
\input{wg_main_software_credit}
\input{wg_main_publishing_SW}
\input{wg_main_user_community}

@@

\input{wg_appendix_best_practices}
\input{wg_appendix_funding_spec_expert}
\input{wg_appendix_catalogs}
\input{wg_appendix_industry_interaction}
\input{wg_appendix_legacy_SW}
\input{wg_appendix_eng_design}
\input{wg_appendix_metrics}
\input{wg_appendix_training}
\input{wg_appendix_software_credit}
\input{wg_appendix_publishing_SW}
\input{wg_appendix_user_community}

>>>>

Turk then moved onto various different definitions of sustainability. His
favorite one was keeping up with bug reports, in which case even if no new
features were added, the software remains sustainable. Another definition of
sustainability Turk mentioned was adding of new features, or maintaining the
software for a long period of time such as the cases of TeX or LaTeX with
community help. A notion Turk heard often at supercomputing conferences was that
sustainable software continued to work on new architectures. Yet another metric
was people continuing to be able to learn how to use and apply the software. A
funder Turk heard talked about sustainability as continuing to get funded. Turk
also recalled that Greg Wilson, among others, said in WSSSPE1.1 that his view of
sustainable software was software that continued to give the same results over
time. A last measure of sustainability Turk presented was the ability to
transition between different people developing and using a piece of software.

At WSSSPE1, several models were presented for ensuring sustainability. Turk
considered that a familiar one was a funded piece of software where an external
agency provided funds to a group that was not necessarily exclusively working on
and developing a piece of software, keeps it going, and provides that out to the
scientific community. The model of productized software, in which a piece of
software has grown into something that individuals rely on to the point that
research groups or people are willing to support it at some pay rate, for
instance, subscription to use cloud services that deploy a piece of software, or
purchase of a piece of software. A final model Turk felt conflicted about is a
volunteer model that is traditional old-school---not modern-day open
source---developement.

Turk discussed whether productizing scientific software was synonoymous with
being sustsainable and self-sufficient. He thought it was not necessarily the
case and furthermore, it could lead to a divergence of interest between users
and developers.

The volunteer model meant unpaid labor. On this note, Turk recommended Ashe
Dryden's blog
post\footnote{http://www.ashedryden.com/blog/the-ethics-of-unpaid-labor-and-the-oss-community}
on the ethics of unpaid labor and the open source software community, which
mentions GitHub as one's resume, for example. Often times, a person funded to
work on some scientific project full time has a small amount of time for working
on a piece of software necessary for the project. However, researchers's ability
to participate in that volunteer community is not always the same and it is not
always aligned with their research projects. From Turk's experience, we cannot
always rely on unpaid labor and volunteer time to sustain a piece of software---
that came down to the notions of the top-down and the bottom-up approaches, the
funded versus the grassroot.

@@

\input{wg_main_best_practices}
\input{wg_main_funding_spec_expert}
%\input{wg_main_catalogs}
\input{wg_main_industry_interaction}
\input{wg_main_legacy_SW}
\input{wg_main_eng_design}
\input{wg_main_metrics}
\input{wg_main_training}
\input{wg_main_software_credit}
\input{wg_main_publishing_SW}
\input{wg_main_user_community}

@@

\input{wg_appendix_best_practices}
\input{wg_appendix_funding_spec_expert}
%\input{wg_appendix_catalogs}
\input{wg_appendix_industry_interaction}
\input{wg_appendix_legacy_SW}
\input{wg_appendix_eng_design}
\input{wg_appendix_metrics}
\input{wg_appendix_training}
\input{wg_appendix_software_credit}
\input{wg_appendix_publishing_SW}
\input{wg_appendix_user_community}
